---
title: "Project1: Lyrics Analysis"
author: "Qing Gao"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytext)
library(dplyr)
library(wordcloud)
library(topicmodels)
library(wesanderson)
library(tm)
```

This report focuses on the analysis of the song lyrics and the differences between several music genre. 

```{r someVar, echo=FALSE}
# load processed lyrics data & artists data
load('../output/processed_lyrics.RData') 
artists.csv <- read.csv('../data/artists.csv')
```


## Explore the Data
```{r, echo=FALSE}
# select the genre from the processed lyrics
genre <- dt_lyrics %>% select(genre, stemmedwords)

# rank of the genre
rank <- genre %>% group_by(genre) %>%
  count() %>%
  arrange(desc(n))

# frequency plot of the genre
ggplot(genre, aes(x = reorder(genre, genre, function(x)-length(x)))) +
  geom_bar() +
  labs(x = 'Genre', y = 'Frequency',
       title = 'Genre Frequency') +
  theme_light()

# year
year <- dt_lyrics %>%
  group_by(year) %>%
  count() %>%
  arrange(year) %>%
  filter(year > 1900)

ggplot(year, aes(x = year,y = n)) +
  geom_bar(stat = 'identity') +
  labs(x = 'Year', y = 'Frequency',
       title = 'Year Frequency') +
  theme_light()

# Year: 2006
year_06 <- dt_lyrics %>% filter(year == 2006) %>%
  group_by(artist) %>%
  count()

```

First, we need to take a look of our data and explore some features of it. From the genre frequency plot, we can find that Rock, Pop and Metal music data dominants our datasets, and we will explore more about these major music genre later and find out their differences. 

The year frequency plot shows that the lyrics data mainly covers from 1970 to 2016. There is more data from year 2006 and 2007. 

## Sentiment Analysis
Since we have more information on rock, pop and metal music, we will compare between these three music genre with the tool of sentiment analysis. In the sentiment anlaysis, we use the package `get_sentiments("nrc")` to assign each word with an emotion from the following ten emotions, which are positive, negative, joy, sadness, fear, anticipation, trust, anger, disgust and surprise. 

### Genre & Emotion: Rock
Let's look at Rock music at first, and do a sentiment analysis on its lyrics. 
```{r, echo=FALSE}
# get sentiment for lyrics in the rock music
rock_lyrics <- dt_lyrics %>% filter(genre == 'Rock') %>% 
  select(stemmedwords) %>%
  unnest_tokens(word, stemmedwords) %>%
  left_join(get_sentiments("nrc"), by = c('word')) %>%
  filter(!is.na(sentiment))

# sentiment percent for rock music
rock_sentiment <- rock_lyrics %>% group_by(sentiment) %>%
  count() %>%
  arrange(desc(n)) %>%
  ungroup() %>%
  mutate(n = n / sum(n)) %>%
  rename(percent = n)
rock_sentiment

# drawing a pie chart
ggplot(rock_sentiment, aes(x = '', y = percent, fill = sentiment)) +
  geom_bar(width = 1, stat = 'identity') +
  coord_polar('y', start = 0) +
  labs(x = '', y = '', title = 'Sentiments for Rock Music') +
  scale_fill_brewer(palette = "Set3") +
  theme_light()

```

Rock music has the positive and negative sentiments the most, which are 17% and 16% of the total sentiments respectively. We can see that most rock music has the positive and joy sentiment. 

```{r, echo=FALSE}
# select high frequency words from each emotions
top10_rock <- rock_lyrics %>%
  count(word, sentiment) %>%
  group_by(sentiment) %>%
  top_n(n = 10) %>%
  ungroup()

# plot the high frequency words with group of different emotions
ggplot(top10_rock, aes(x = reorder(word, n), y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales="free") +
  labs(x = 'Word', y = 'Contribution to sentiment') +
  coord_flip() +
  theme_light()

# Word Cloud: Rock
pal = brewer.pal(5, "Set3")
rock_lyrics %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100, colors = rev(pal)))
```

We look closer to the details of rock music's lyrics. The most frequent word for positive sentiment is love and for negative one is ill. From the word cloud, we can also see that love and ill appear the most frequent in the rock music. 


### Genre & Emotion: Pop
Let's look at the pop music. 
```{r, echo=FALSE}
# get sentiment for lyrics in the rock music
pop_lyrics <- dt_lyrics %>% filter(genre == 'Pop') %>% 
  select(stemmedwords) %>%
  unnest_tokens(word, stemmedwords) %>%
  left_join(get_sentiments("nrc"), by = c('word')) %>%
  filter(!is.na(sentiment))

# sentiment percent for rock music
pop_sentiment <- pop_lyrics %>% group_by(sentiment) %>%
  count() %>%
  arrange(desc(n)) %>%
  ungroup() %>%
  mutate(n = n / sum(n)) %>%
  rename(percent = n)
pop_sentiment

# drawing a pie chart
ggplot(pop_sentiment, aes(x = '', y = percent, fill = sentiment)) +
  geom_bar(width = 1, stat = 'identity') +
  coord_polar('y', start = 0) +
  labs(x = '', y = '', title = 'Sentiments for Pop Music') +
  scale_fill_brewer(palette="Paired") +
  theme_light()


```

According to the table and pie chart, pop music has more positive sentiment than that of rock music. There are 34% of sentiment are positive and joy in pop music lyrics.  

```{r, echo=FALSE}
# select high frequency words from each emotions
top10_pop <- pop_lyrics %>%
  count(word, sentiment) %>%
  group_by(sentiment) %>%
  top_n(n = 10) %>%
  ungroup()

# plot the high frequency words with group of different emotions
ggplot(top10_pop, aes(x = reorder(word, n), y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales="free") +
  labs(x = 'Word', y = 'Contribution to sentiment') +
  coord_flip() +
  theme_light()

# Word Cloud: Pop
pal = brewer.pal(6, "Paired")
pop_lyrics %>% anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100, colors = rev(pal)))
```

From the detailed analysis of the pop music lyrics, "love", "feeling" and 'baby' dominates the lyrics. In short conclusion, pop music has more positive and joy sentiments 

### Genre & Emotion: Metal
```{r, echo=FALSE}
# get sentiment for lyrics in the rock music
metal_lyrics <- dt_lyrics %>% filter(genre == 'Metal') %>% 
  select(stemmedwords) %>%
  unnest_tokens(word, stemmedwords) %>%
  left_join(get_sentiments("nrc"), by = c('word')) %>%
  filter(!is.na(sentiment))

# sentiment percent for rock music
metal_sentiment <- metal_lyrics %>% group_by(sentiment) %>%
  count() %>%
  arrange(desc(n)) %>%
  ungroup() %>%
  mutate(n = n / sum(n)) %>%
  rename(percent = n)
metal_sentiment

# drawing a pie chart
ggplot(metal_sentiment, aes(x = '', y = percent, fill = sentiment)) +
  geom_bar(width = 1, stat = 'identity') +
  coord_polar('y', start = 0) +
  labs(x = '', y = '', title = 'Sentiments for Metal Music') +
  scale_fill_brewer(palette="RdGy") +
  theme_light()
```

Metal music is different from both rock and pop music. It has more negative, fear, sadness and anger sentiments in the lyrics. Most of its lyrics have negative emotions.


```{r, echo=FALSE}
# select high frequency words from each emotions
top10_metal <- metal_lyrics %>%
  count(word, sentiment) %>%
  group_by(sentiment) %>%
  top_n(n = 10) %>%
  ungroup()

# plot the high frequency words with group of different emotions
ggplot(top10_metal, aes(x = reorder(word, n), y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales="free") +
  labs(x = 'Word', y = 'Contribution to sentiment') +
  coord_flip() +
  theme_light()

# Word Cloud: Metal
pal = brewer.pal(5, "RdGy")
metal_lyrics %>% anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100, colors = rev(pal)))
```

The words with high frequency in metal music are 'death', 'ill', and 'die'. We can conclude that metal music has more negative sentiments than other two music genre. 


## Topic Modeling
### Rock
```{r, echo=FALSE}
# create document-term matrix of rock lyrics
dt_sample <- dt_lyrics %>% filter(genre == 'Rock') %>% sample_n(2000)
corpus <- Corpus(VectorSource(c(dt_sample$stemmedwords)))
dtm <- DocumentTermMatrix(corpus)

#Set parameters for Gibbs sampling
burnin <- 4000
iter <- 2000
thin <- 500
seed <-list(2003,5,63,100001,765)
nstart <- 5
best <- TRUE
#Number of topics
k <- 5
#Run LDA using Gibbs sampling
ldaOut <-LDA(dtm, k, method="Gibbs", control=list(nstart=nstart, 
                                                 seed = seed, best=best,
                                                 burnin = burnin, iter = iter, 
                                                 thin=thin))
#write out results
#docs to topics
ldaOut.topics_rock <- as.matrix(topics(ldaOut))
table(c(1:k, ldaOut.topics_rock))

#top 6 terms in each topic
ldaOut.terms_rock <- as.matrix(terms(ldaOut,6))
ldaOut.terms_rock

#probabilities associated with each topic assignment
topicProbabilities_rock <- as.data.frame(ldaOut@gamma)

```

### Pop
```{r, echo=FALSE}
# create document-term matrix of rock lyrics
dt_sample <- dt_lyrics %>% filter(genre == 'Pop') %>% sample_n(2000)
corpus <- Corpus(VectorSource(c(dt_sample$stemmedwords)))
dtm <- DocumentTermMatrix(corpus)

#Set parameters for Gibbs sampling
burnin <- 4000
iter <- 2000
thin <- 500
seed <-list(2003,5,63,100001,765)
nstart <- 5
best <- TRUE
#Number of topics
k <- 5
#Run LDA using Gibbs sampling
ldaOut <-LDA(dtm, k, method="Gibbs", control=list(nstart=nstart, 
                                                 seed = seed, best=best,
                                                 burnin = burnin, iter = iter, 
                                                 thin=thin))
#write out results
#docs to topics
ldaOut.topics_pop <- as.matrix(topics(ldaOut))
table(c(1:k, ldaOut.topics_pop))

#top 6 terms in each topic
ldaOut.terms_pop <- as.matrix(terms(ldaOut,6))
ldaOut.terms_pop

#probabilities associated with each topic assignment
topicProbabilities_pop <- as.data.frame(ldaOut@gamma)

```

### Metal
```{r, echo=FALSE}
# create document-term matrix of rock lyrics
dt_sample <- dt_lyrics %>% filter(genre == 'Metal') %>% sample_n(2000)
corpus <- Corpus(VectorSource(c(dt_sample$stemmedwords)))
dtm <- DocumentTermMatrix(corpus)

#Set parameters for Gibbs sampling
burnin <- 4000
iter <- 2000
thin <- 500
seed <-list(2003,5,63,100001,765)
nstart <- 5
best <- TRUE
#Number of topics
k <- 5
#Run LDA using Gibbs sampling
ldaOut <-LDA(dtm, k, method="Gibbs", control=list(nstart=nstart, 
                                                 seed = seed, best=best,
                                                 burnin = burnin, iter = iter, 
                                                 thin=thin))
#write out results
#docs to topics
ldaOut.topics_metal <- as.matrix(topics(ldaOut))
table(c(1:k, ldaOut.topics_metal))

#top 6 terms in each topic
ldaOut.terms_metal <- as.matrix(terms(ldaOut,6))
ldaOut.terms_metal

#probabilities associated with each topic assignment
topicProbabilities_metal <- as.data.frame(ldaOut@gamma)
```

The topic modeling shows the different five topics under the music genre rock, pop and metal. 

## Conclusion
The dataset lyrics contains 100,000+ lyrics information. Since the dataset contains overwhelming information on the lyrics, this analysis focuses on the comparison between several music genre, which are rock, pop and metal music.

From the sentiment analysis, rock and pop music have more positive and joy sentiments than the metal music. While lyrics in metal music has a lot of negative and sadness sentiments. 

From the topic modeling, we find out five topics and top 6 terms for each music genre respectively.  

